---
title: "Introduction to Spatial Data Analysis in R"
subtitle: "useR! 2022"
author: "Adithi Upadhya, Meenakshi Kushwaha"
institute: "ILK Labs"
date: "20th June 2022 (updated: `r Sys.Date()`)"
output:
  html_document:
    code_folding: hide
---

### Load Libraries  


```{r, message = FALSE, warning = FALSE}
library(sf)
library(raster)
library(tidyverse)
library(stars)
library(leaflet)
library(ggplot2)
library(ggspatial)
```


### st_as_sf

Convert foreign object to an `sf` object.


```{r, message = FALSE, warning = FALSE}
data <- read.csv("data/spotted.csv", sep = ",")
prj4string <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
my_projection <- st_crs(prj4string)
class(data)

# Create sf object
data_sf <- st_as_sf(data, coords = c("Longitude", "Latitude"), crs = my_projection)
class(data_sf)
plot(data_sf)
```


### Map


```{r message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = data_sf) + 
  coord_sf()
```


### st_transform

Transform or convert coordinates of simple feature.

```{r, message = FALSE, warning = FALSE}
data_sf_projected <- st_transform(data_sf, 32643)
data_sf_projected
```


### st_buffer

Computes a buffer around this geometry/each geometry.

```{r, message = FALSE, warning = FALSE}
buf <- st_buffer(data_sf_projected, dist = 3000)
plot(buf)
```


### st_intersects

Returns `TRUE` or `FALSE.`

```{r, message = FALSE, warning = FALSE}
st_intersects(data_sf_projected[1, ], data_sf_projected[2, ])
st_intersects(data_sf_projected[1, ], data_sf_projected[1, ])
```


### st_join

```{r message = FALSE, warning = FALSE}
data_join <- read.csv("data/spotted_join.csv", sep = ",")
# Create sf object
data_join_sf <- st_as_sf(data_join, coords = c("Longitude", "Latitude"), crs = my_projection)

# Join using predicate 
joined <- st_join(data_sf, data_join_sf, join = st_equals)
joined
```



### Raster 


```{r, message = FALSE, warning = FALSE}
temp <- raster("data/Ban_Temp_Mar2022.tif")
temp
```


### projectRaster

To transform a `RasterLayer` to another coordinate reference system (projection)

```{r, message = FALSE, warning = FALSE}
# To transform 
temp_proj <- projectRaster(temp, crs = "+proj=utm +zone=43 +datum=WGS84")
```


### Mask by buffer 

Mask values from a `Raster` object at the locations of spatial vector data

```{r, message = FALSE, warning = FALSE}
data_sf_subset <- data_sf[1, ]

data_sf_buffer <- st_buffer(data_sf_subset, dist = 12000) 
masked_raster <- mask(x = temp, mask = data_sf_buffer)
```


### Calculate area

```{r, message = FALSE, warning = FALSE}
sum(getValues(area(temp_proj)))
```


### Rasterize

```{r, message = FALSE, warning = FALSE}
raster_data <- st_rasterize(data_sf)
class(raster_data)
plot(raster_data)
```


### Vectorize 
 
```{r, message = FALSE, warning = FALSE}
tif <- read_stars("data/Ban_Temp_Mar2022.tif")
class(tif)
# Convert a foreign object to sf
temp_sf <- st_as_sf(tif, as_points = TRUE)
class(temp_sf)
plot(temp_sf)
```

### Static Map

```{r, message = FALSE, warning = FALSE}
bangalore_boundary <- st_read("data/BBMP_Boundary.shp")
bangalore_boundary <- st_transform(bangalore_boundary, crs = my_projection)

ggplot() + geom_sf(data = bangalore_boundary, color = "green", size = 2, alpha = 0.1, fill = "green") +
  geom_sf(data = data_sf, color = "blue", size = 2) + theme_bw() +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  coord_sf(xlim = c(77.4, 77.8), ylim = c(12.8, 13.2))
```

# Interactive Map

```{r, message = FALSE, warning = FALSE}
leaflet(bangalore_boundary) %>% 
  addTiles() %>% 
  addPolygons(color = "green") %>%
      addCircles(
        data = data,
        lng = ~ Longitude,
        lat = ~ Latitude,
        radius = 20,
        weight = 20,
        fill = TRUE)
```
